Supplementary Materials to ``Statistical paradoxes in coronavirus case-counts: selection bias, measurement error, and the COVID-19 pandemic''

### 2022-09-12 write-up/letters/response-ref-1-v1.tex

%----------------------------------------------------------------------------------------
%	LETTER CONTENT SECTION
%----------------------------------------------------------------------------------------

\vspace{5mm}

\newpage

{\bf Response to Referee}

We appreciate the very helpful comments from the referee which has helped improve the paper substantially. We have addressed the concerns as best as possible in this revision. We provide a further point-by-point explanation below. Any comments in {\it italics} indicate a statement from the referee report.
\begin{enumerate}
\item {\it I appreciate the addition of the remark that states that the dynamics mentioned in this paper refer to the original strain. There are still lines (for example: “Upon infection with SARS-CoV-2, an incubation period (time to symptom onset) starts and lasts approximately five days”) that are more definitive than I would prefer. Perhaps adding in “Upon infection with the original SARS-CoV-2 variant” would be preferable.}

\vspace{5mm}
We agree with this point and have gone through the manuscript to reduce the chances a reader miscontrues our explanations. We aim for clarity in these edits by avoiding definitive statements while not making statements that are too vauge or ambiguous.
\vspace{5mm}

\item {\it The reliance on the random sample with such a low response rate may be subject to unmeasured confounding - perhaps a sensitivity analysis showing the impact of a potential unmeasured confounder could strengthen this point.}

\vspace{5mm}
We have added a sensitivity analysis to the supplementary materials where we consider an unmeasured binary confounder~$Z_{j,t} \in \{0,1\}$.  In this setting, {\it prevalence}~$pr(Z=1|X,Y)$ and {\it propensity} $\pi(X, Z; (\theta,\theta_z))$ both impact the sensitivity analysis.  Imputation-based sensitivity analyses will thus depend heavily on the choice of prevalence and $\theta_z$.    Therefore, we consider a Taylor series expansion of the IPW and DR-estimators at the estimated propesity which is equivalent to~$\theta_z = 0$.denote the coefficient in the selection propensity model and
$$
\hat \mu_{IPW} + \theta_z \frac{1}{\sum_{j=1}^N I^{NR}_j w_j (X_{j,t}; \hat \theta_t)} \left[ \sum_{j=1}^N I^{NR}_j \frac{1-\pi_j(X_{j,t}; \hat \theta_t)}{\pi_j(X_{j,t}; \hat \theta_t)} E[Z_j] (Y_j - \hat \mu_{IPW}) \right]
$$
Thus bias depends on the correlation of the error $Y_j - \hat \mu_{IPW}$ to the odds ratio times the expected value of the unmeasured binary confounder.  We use this to calculate the

\item {\it On the plots that show the active infection rate estimates, perhaps including a line that is the case counts divided by the population as a floor would be worthwhile (at least in the supplement) to demonstrate that the model-based estimates are at least above this.}

\vspace{5mm}
We have added this this floor .  We clarify that ``case counts'' in the comment above should be a measure of recent case counts to be a relevant  active infection rate
\vspace{5mm}

\item {\it I found the grey hard to differentiate for the figures that have more than two categories – perhaps varying the shapes would help if trying to keep these grey-scale?}

\vspace{5mm}
We have added shapes to help differentiate figures while keeping the grey-scale.
\vspace{5mm}

### 2022-09-12 write-up/covid-supplement-writeup.tex
Update covid-supplement-writeup.tex

The proposed method relies on the assumption $I_{j,t}^{(NR)} \indep Y_{j,t} \, | \, X_{j,t}$, i.e., the sampling indicator in the non-probabilistic survey is conditionally independent of the outcome given the covariates.  However, estimates of population quantities based on inverse-weighting will be biased in the presence of `unobserved confounding'.  We may hope that the bias

The purpose of this section is to introduce XXX,

We consider a
\begin{align*}
\logit \pi \left( I_{t} = 1  | x_{t}, u_{t} \right) &= h_t(x_t) + \alpha u \\
E[ Y | i_t, x_t, u_t] = l(i_t, x_t)
\end{align*}
for some function~$h_t$, $l_t$ that are allowed to depend on time. Re-arranging the terms we see that .  This can be
$$
E[ Y | i_t x_t, u_t ] = l(t,x)  + \frac{\delta}{\alpha} \left( \logit \pi \left( I_{t} = 1  | x_{t}, u_{t} \right) - h_t (x_t) \right)
$$
The key idea in XXXX is that positing a distribution on $P(I_t = 1 | x_y, u_t)$ directly allows one to circumvent the need to specify a distribution for~$U_t$, a highly non-trivial task.

>>>>

The proposed method relies on the assumption $I_{j,t}^{(NR)} \indep Y_{j,t} \, | \, X_{j,t}$, i.e., the sampling indicator in the non-probabilistic survey is conditionally independent of the outcome given the covariates.  However, estimates of population quantities based on inverse-weighting and/or model-based estimates will be biased in the presence of `unobserved confounding'.  The purpose of this section is to introduce a sensitivity analysis.  To this, we consider the logit-linear model by XXXX adjusted to self-selection rather than causal inference.  That is,
\begin{align*}
\logit \pi \left( I_{t} = 1  | x_{t}, u_{t} \right) &= h_t(x_t) + \alpha u \\
E[ Y | i_t, x_t, u_t] &= l(i_t, x_t) + \delta u
\end{align*}
for some function~$h_t$, $l_t$ that are allowed to depend on time. Re-arranging the terms we see that .  This can be
$$
E[ Y | i_t x_t, u_t ] = l(t,x)  + \frac{\delta}{\alpha} \left( \logit \pi \left( I_{t} = 1  | x_{t}, u_{t} \right) - h_t (x_t) \right)
$$
The key idea in XXXX is that positing a distribution on $P(I_t = 1 | x_y, u_t)$ directly allows one to circumvent the need to specify a distribution for~$U_t$, a highly non-trivial task.
